---
title: "finalProjectPipkin"
author: "Daniel Pipkin"
date: "5/6/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Setting up the data

library(tidyverse)
HilmanDataCSV = read.csv(url("https://raw.githubusercontent.com/dpipkin99/IS4300-Final-Project/master/HilmanData.csv"))
HD<- as.data.frame(HilmanDataCSV)

#Cleaning the Data
  #Classification
HD$Classification<-factor(HD$Classification, levels = c("Freshman", "Sophomore", "Junior","Senior","Graduate"))
  #Socio-Politico-Economic Leaning
HD$PoliticalLean <-factor(HD$PoliticalLean, levels=c("Very Liberal", "Somewhat Liberal", "Independent", "Somewhat Conservative", "Very Conservative"))
HD$SocialLean <-factor(HD$SocialLean, levels=c("Very Liberal", "Somewhat Liberal", "Independent", "Somewhat Conservative", "Very Conservative"))
HD$EconomicLean <-factor(HD$EconomicLean, levels=c("Very Liberal", "Somewhat Liberal", "Independent", "Somewhat Conservative", "Very Conservative"))
  #Initial News Opinions
HD$WaPoCredible<-factor(HD$WaPoCredible, levels=c("Strongly agree", "Somewhat agree","Somewhat disagree","Strongly disagree"))
HD$WaPoWhole<-factor(HD$WaPoWhole, levels=c("Strongly agree", "Somewhat agree","Somewhat disagree","Strongly disagree"))
HD$FoxCredible<-factor(HD$FoxCredible, levels=c("Strongly agree", "Somewhat agree","Somewhat disagree","Strongly disagree"))
HD$FoxWhole<-factor(HD$FoxWhole, levels=c("Strongly agree", "Somewhat agree","Somewhat disagree","Strongly disagree"))

    #Converting To Factor
HD$WaPoRate1<-factor(HD$WaPoRate1, levels=c("True", "Mostly True", "Mostly False", "False"))
HD$WaPoRate2<-factor(HD$WaPoRate2, levels=c("True", "Mostly True", "Mostly False", "False"))
HD$FoxRate1<-factor(HD$FoxRate1, levels=c("True", "Mostly True", "Mostly False", "False"))
HD$FoxRate2<-factor(HD$FoxRate2, levels=c("True", "Mostly True", "Mostly False", "False"))

  #Washington Post Question 1
HD$WaPoCheckOther1 <- factor(HD$WaPoCheckOther1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoAskOthers1 <- factor(HD$WaPoAskOthers1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoSocialMed1 <- factor(HD$WaPoSocialMed1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoSearchOnline1 <- factor(HD$WaPoSearchOnline1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoOtherSource1 <- factor(HD$WaPoOtherSource1,levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
  #Washington Post Question 2
HD$WaPoCheckOther2 <- factor(HD$WaPoCheckOther2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoAskOthers2 <- factor(HD$WaPoAskOthers2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoSocialMed2 <- factor(HD$WaPoSocialMed2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoSearchOnline2 <- factor(HD$WaPoSearchOnline2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$WaPoOtherSource2 <- factor(HD$WaPoOtherSource2,levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
  #Fox News Question 1
HD$FoxCheckOther1 <- factor(HD$FoxCheckOther1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxAskOthers1 <- factor(HD$FoxAskOthers1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxSocialMedia1 <- factor(HD$FoxSocialMedia1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxSearchOnline1 <- factor(HD$FoxSearchOnline1, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxOtherSource1 <- factor(HD$FoxOtherSource1,levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
  #Fox News Question 2
HD$FoxCheckOther2 <- factor(HD$FoxCheckOther2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxAskOthers2 <- factor(HD$FoxAskOthers2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxSocialMedia2 <- factor(HD$FoxSocialMedia2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxSearchOnline2 <- factor(HD$FoxSearchOnline2, levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"))
HD$FoxOtherSource2 <- factor(HD$FoxOtherSource2,levels= c("Very Likely","Somewhat Likely", "Somewhat Unlikely", "Very Unlikely"));



  #Statistics
#Generating a general feel for the overall lean of the respondents
#With 1 being very liberal, 5 being very conservative, and 3 being more independent.
meanPolitical = mean(as.numeric(HD$PoliticalLean))
meanSocial = mean(as.numeric(HD$SocialLean))
meanEconomical = mean(as.numeric(HD$EconomicLean))
#Creating an aggregate lean
overallMean = ((meanPolitical+meanSocial+meanEconomical)/3)
#Overall mean is 3.48. Slight lean conservative.
#Creating individual means for each student
HD<-HD %>% mutate(indivMean = (as.numeric(PoliticalLean)+as.numeric(SocialLean)+as.numeric(EconomicLean))/3)
#Seperating into generally conservative and generally liberal
HDCon<-HD%>%filter(indivMean>3.5)
HDLib<-HD%>%filter(indivMean<3.5)
#Creating GGPlots of each data subset
HDPlot <- HD%>%ggplot()
HDPlotCon <- HDCon%>%ggplot()
HDPlotLib <- HDLib%>%ggplot()

```

## Introduction

This report is based on research conducted by Justin Hilman. For a journalism research project, Justin shared a poll with Harding students to learn about their news consumption habits.  

Justin asked questions about people's social, political, and economic beliefs to get a basis of people's biases.  

Justin then asked questions about whether people think that the Washington Post and Fox News are credible and whether they tell the whole story.  

After getting preliminary information about people's biases, Justin asks for the student's opinion on a headline and whether they would verify the headline using an outside source. This was done a total of four times. With two headlines per news outlet. One headline was an actual headline, the other was one that was fake.  

The variables that are used from the data set are as follows:

*  PoliticalLean: Factor, The political stance of the respondent (Very Liberal,Somewhat Liberal, Independent, Somewhat Conservative, Very Conservative)  

*  SocialLean: Factor, The social stance of the respondent (Very Liberal,Somewhat Liberal, Independent, Somewhat Conservative, Very Conservative)  

*  EconomicLean: Factor, The economic stance of the respondent (Very Liberal,Somewhat Liberal, Independent, Somewhat Conservative, Very Conservative)  

*   WaPoCredible: Factor, How credible the respondent thinks the Washington Post is (Strongly agree, Somewhat agree, Somewhat disagree, Strongly disagree)  

*   WaPoWhole: Factor, How much of the story the respondent thinks The Washington Post tells. (Strongly agree, Somewhat agree, Somewhat disagree, Strongly disagree)  

*   FoxCredible: Factor, How credible the respondent thinks Fox News is.  (Strongly agree, Somewhat agree, Somewhat disagree, Strongly disagree)  

*   FoxWhole: Factor, How much of the story the respondent thinks Fox News tells.  (Strongly agree, Somewhat agree, Somewhat disagree, Strongly disagree)  
  
  

Other variables used:

*   meanPolitical: Numeric, The mean of PoliticalLean for the data set.

*   meanSocial: Numeric, The mean of SocialLean for the data set.  
*   meanEconomic: Numeric, The mean of EconomicLean for the data set.

*   overallMean: Numeric, a value generated by taking meanPolitical, meanSocial, and meanEconomic, used to generate an overall feel for a mean lean of the data set.  

*   indivMean: Numeric, the mean of meanPolitical, meanSocial, and meanEconomic per respondent. Used to generate an overall political bias.  

*  HDCon: Dataset filtered with indivMean values greater than 3.5.  
*  HDLib: Dataset filtered with indivMean values less than 2.5.  

The following section of variables are based off of the following 4 questions:  
  
Washington Post:  

1.  "How would you rate the following Washington Post Headline: "President Biden vows to dedicate four trillion dollars to climate change science""  

2.  "How would you rate the following Washington Post headline: "Biden's messed-up math comparing war deaths to covid deaths""  
  
Fox News:  

1.  "How would you rate the following Fox News headline: "Biden makes statement claiming that all transgender women should be able to compete in women's sports""  

2.  "How would you rate the following Fox News headline: "Biden 'optimistic' about bipartisan work to 'end cancer as we know it'""  

The variables that correspond with the questions are as folllows:

*  WaPoRate[1/2]: Factor, the respondent's rating of how true they believe the headline to be.
*  FoxRate[1/2]: Factor, the respondent's rating of how true they believe the headline to be.
  


```{r, include=FALSE}

#Less than .05. Reject the Null that the mean is less than 2. 
WaPoQ1 <- t.test(as.numeric(HDCon$WaPoRate1), mu = 2, alternative = "greater")
#Greater than .05 accept the null that the mean is less than 2.
FoxQ1 <- t.test(as.numeric(HDCon$FoxRate1), mu = 2, alternative = "greater")
#P value .1121
WaPoQ2 <- t.test(as.numeric(HDCon$WaPoRate2), mu = 2, alternative = "greater")
# Reject the null since 
FoxQ2 <- t.test(as.numeric(HDCon$FoxRate2), mu = 2, alternative = "greater")


```

## Including Plots

You can also embed plots, for example:

```{r, echo = FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
